Simple repression draft

### 2020-09-08 doc/resubmission/appendix_01_nonbursty.tex
first attempt to shrink section 02. Sending to appendix 01 in the mean time'

@@

\begin{figure}[h!]
\centering
\includegraphics{../figures/si/figS0X_two_state.pdf}
\caption{
\textbf{Two-state promoter chemical master equation.}
(A) Schematic of the two state promoter simple repression model. Rates $k_R^+$
and $k_R^-$ are the association and dissociation rates of the transcriptional
repressor, respectively, $r$ is the transcription initiation rate, and $\gamma$
is the mRNA degradation rate. (B) Schematic depiction of the mRNA count state
transitions. The model in (A) only allows for jumps in mRNA of size 1. The
production of mRNA can only occur when the promoter is in the transcriptionally
active state.}
\label{fig:two_state}
\end{figure}

>>>>

\mrm{THIS NEEDS TO BE FIXED}
\subsubsection{The Two-State Equilibrium Model}
In this simplest model, depicted as (1) in Figure~\ref{fig1:means_cartoons}(B),
the promoter is idealized as existing in one of two states, either repressor
bound or repressor unbound. The rate of transcription
is assumed to be proportional to the fraction of time spent
in the repressor unbound state.
From the relative statistical weights listed in
Figure~\ref{fig1:means_cartoons}, the probability $p_U$ of being in the
unbound state is
\begin{equation}
p_U = \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}\right)^{-1}.
\end{equation}
The mean rate of transcription is then given by $r p_U$, as assumed by
Eq.~\ref{eq:transcrip_prop_pbound}. The mean number of mRNA is
set by the balance of average mRNA transcription and degradation rates,
so it follows that the mean mRNA level is given by
\begin{equation}
\langle m \rangle = \frac{r}{\gamma}
        \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}\right)^{-1},
\end{equation}
where $r$ is the transcription rate from the repressor unbound state, $\gamma$
is the mRNA degradation rate, $R$ is repressor copy number, $N_{NS}$ is the
number of nonspecific binding sites in the genome where repressors spend most
of their time when not bound to the operator, $\beta \equiv 1/k_BT$, and
$\Delta\varepsilon_R$ is the binding energy of a repressor to its operator site.
The derivation of this result is deferred to Appendix~\ref{sec:non_bursty}.

The fold-change is  found as the ratio of mean mRNA with and without repressor
as introduced in Eq.~\ref{eq:fc_def}. Invoking that definition results in
\begin{equation}
\text{fold-change}
= \left(1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}\right)^{-1},
\end{equation}
which matches the form of the master curve in
Figure~\ref{fig1:means_cartoons}(D) with $\rho=1$ and $\Delta F_R = 
\beta\Delta\varepsilon_r - \log (R / N_{NS})$.

In fact it was noted in~\cite{Chure2019} that this two-state model can be viewed
as the coarse-graining of any equilibrium promoter model in which no
transcriptionally active states have transcription factor bound, or put
differently, when there is no overlap between
transcription factor bound states and transcriptionally
active states. We will see this explicitly in the 3-state equilibrium model
below, but perhaps surprising is that an analogous result carries over even to
the nonequilibrium models we consider later.

\subsubsection{The Three-State Equilibrium Model}
Compared to the previous model, here we fine-grain the repressor unbound state
into two separate states: empty, and RNAP bound as shown in (2) in
Figure~\ref{fig1:means_cartoons}(B). This picture was used in~\cite{Garcia2011a}
as we use it here, and in~\cite{Razo-Mejia2018} and~\cite{Chure2019} it was
generalized to incorporate small-molecule inducers that bind the repressor. 
The effect of this generalization is, roughly speaking,
simply to rescale $R$ from the total number of repressors to a smaller
effective number of available repressors which are unbound by inducers.
We point out that the same generalization can be incorporated quite easily
into any of our models in Figure~\ref{fig1:means_cartoons} by simply
rescaling the repressor copy number $R$ in the equilibrium models, or
equivalently $k_R^+$ in the nonequilibrium models.

The mean mRNA copy number, as derived in Appendix~\ref{sec:non_bursty}
from a similar enumeration of states and weights as the previous model, is
\begin{equation}
\langle m \rangle = \frac{r}{\gamma}
\frac{\frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
        {
        1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}
        + \frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}
        },
\end{equation}
where the new variables are $\Delta\varepsilon_P$, the difference in RNAP
binding energy to its specific site (the promoter) relative to an average
nonspecific background site, and the RNAP copy number, $P$. The fold-change
again follows immediately as
\begin{align}
\text{fold-change}
&= \frac{\frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
        {
        1 + \frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}
        + \frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}
        }
\frac{1 + \frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
        {\frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
\\
&= \left(
1 + \frac{\frac{R}{N_{NS}} e^{-\beta\Delta\varepsilon_R}}
        {1 + \frac{P}{N_{NS}} e^{-\beta\Delta\varepsilon_P}}
\right)^{-1}
\\
&= (1 + \exp(-\Delta F_R - \log\rho))^{-1},
\end{align}
with $\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})$ and $\rho = 1 +
\frac{P}{N_{NS}}\mathrm{e}^{-\beta\Delta\varepsilon_P}$ as shown
in~\fig{fig1:means_cartoons}(B). Thus far, we see that the two thermodynamic
models, despite making different coarse-graining commitments, result in the same
functional form for the fold-change in mean gene expression.  We now explore how
kinetic models fare when faced with computing the same observable.
\mrm{ENDS HERE}

\mrm{THIS NEEDS TO BE FIXED}
\subsubsection{The Poisson Promoter Nonequilibrium Model}
For our first kinetic model, we  imitate the states considered in the Two-State
Equilibrium Model and consider the simplest possible picture with only two
states, repressor bound and unbound. This is exactly the model used for the main
results of~\cite{Jones2014}. In this picture, repressor association and
dissociation rates from its operator site, $k_R^+$ and $k_R^-$, respectively,
govern transitions between the two states. When the system is in the unbound
state, transcription initiates at rate $r$, which represents a coarse-graining
of all the downstream processes into a single effective rate. mRNA is degraded
at rate $\gamma$ as already exploited in the previous models.

Let $p_R(m,t)$ denote the joint probability of finding the system in the
repressor bound state $R$ with $m$ mRNA molecules present at time $t$. Similarly
define $p_U(m,t)$ for the repressor unbound state $U$. This model is governed
by coupled master equations giving the time evolution of $p_R(m,t)$ and
$p_U(m,t)$~\cite{Sanchez2008, Sanchez2011, Phillips2019} which we can write as
\begin{align}
\begin{split}
\deriv{t}p_R(m,t) =& 
- \overbrace{k_R^- p_R(m,t)}^{R \rightarrow U}
+ \overbrace{k_R^+ p_U(m,t)}^{U \rightarrow R}
+ \overbrace{(m+1)\gamma p_R(m+1,t)}^{m + 1 \rightarrow m}
- \overbrace{\gamma p_R(m,t)}^{m \rightarrow m - 1}
\\
\deriv{t}p_U(m,t) =&\; 
\overbrace{k_R^- p_R(m,t)}^{R \rightarrow U}
- \overbrace{k_R^+ p_U(m,t)}^{U \rightarrow R}
+ \overbrace{rp_U(m-1,t)}^{m-1 \rightarrow m}
- \overbrace{rp_U(m,t)}^{m \rightarrow m + 1}
\\
&+ \overbrace{(m+1)\gamma p_U(m+1,t)}^{m + 1 \rightarrow m}
- \overbrace{\gamma p_U(m,t)}^{m \rightarrow m - 1},
\label{eq:poisson_promoter_cme}
\end{split}
\end{align}
where each term on the right corresponds to a transition between two states of
the promoter as indicated by the overbrace label. In each equation, the first
two terms describe transitions between promoter states due to repressors
unbinding and binding, respectively. The final two terms describe degradation of
mRNA, decreasing the copy number by one, and the terms with coefficient $r$
describe transcription initiation increasing the mRNA copy number by one.
We direct the reader to Appendix~\ref{sec:cme_from_cartoon} for a careful
treatment showing how the form of this master equation follows from the
corresponding cartoon in Figure~\ref{fig1:means_cartoons}.

We can greatly simplify the notation, which will be especially useful for the
more complicated models yet to come, by re-expressing the master equation in
vector form~\cite{Phillips2012}. The promoter states are collected into a vector
and the rate constants are collected into matrices as
\begin{equation}
\vec{p}(m) = \begin{pmatrix} p_R(m) \\ p_U(m) \end{pmatrix},\
\mathbf{K} = \begin{pmatrix} -k_R^- & k_R^+ \\ k_R^- & -k_R^+ \end{pmatrix},\
\mathbf{R} = \begin{pmatrix} 0 & 0 \\ 0 & r \end{pmatrix},\
\label{eq:2state_cme_matrices}
\end{equation}
so that the master equation may be condensed as
\begin{equation}
\deriv{t}\vec{p}(m,t) =
\left( \mathbf{K} - \mathbf{R} - \gamma m \mathbf{I} \right) \vec{p}(m,t)
                + \mathbf{R} \vec{p}(m-1,t)
                + \gamma (m+1) \mathbf{I} \vec{p}(m+1,t),
\label{eq:2state_rep_cme}
\end{equation}
where $\mathbf{I}$ is the identity matrix. Taking steady state by setting time
derivatives to zero, the mean mRNA can be found to be
\begin{equation}
\langle m \rangle = \frac{r}{\gamma}
        \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1},
\label{eq:mean_m_model1}
\end{equation}
with the algebra details again deferred to Appendix~\ref{sec:non_bursty}. Recall
$k_R^+$ is proportional to the repressor copy number, so in computing
fold-change, absence of repressor corresponds to $k_R^+\rightarrow0$. Therefore
fold-change in this model is simply
\begin{equation}
\text{fold-change} = \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1},
\end{equation}
again matching the master curve of~\fig{fig1:means_cartoons}(D) with $\rho=1$.

\subsubsection{Nonequilibrium Model Two - RNAP Bound and Unbound States}
Our second kinetic model depicted in Figure~\ref{fig1:means_cartoons}(C) mirrors
the second equilibrium model of Figure~\ref{fig1:means_cartoons}(B) by
fine-graining  the repressor unbound state of nonequilibrium model 1, resolving
it into an empty promoter state and an RNAP-bound state. Note in this picture,
in contrast with model 4 below, transcription initiation is accompanied by a
promoter state change, in keeping with the interpretation as RNAP-bound and
empty states: if an RNAP successfully escapes the promoter and proceeds to
elongation of a transcript, clearly it is no longer bound at the promoter.
Therefore another RNAP must bind before another transcript can be initiated.

The master equation governing this model is analogous
to Eqs.~\ref{eq:2state_cme_matrices}-\ref{eq:2state_rep_cme} for model 1 above. The
main subtlety arises since transcription initiation accompanies a promoter state
change. This can be understood by analogy to $\mathbf{K}$. The off-diagonal and
diagonal elements of $\mathbf{K}$ correspond to transitions arriving at or
departing from, respectively, the promoter state of interest. If transcription
initiation is accompanied by promoter state changes, we must have separate
matrices for arriving and departing transcription events since the arriving and
departing transitions have different initial copy numbers of mRNA, unlike for
$\mathbf{K}$ where they are the same (see Appendix~\ref{sec:non_bursty}). The
master equation for this model is
\begin{equation}
\deriv{t}\vec{p}(m,t) =
\left( \mathbf{K} - \mathbf{R_D} - \gamma m \mathbf{I} \right) \vec{p}(m,t)
                + \mathbf{R_A} \vec{p}(m-1,t) +
                \gamma (m+1) \mathbf{I} \vec{p}(m+1,t),
\label{eq:3state_rep_cme}
\end{equation}
with the state vector and promoter transition matrix defined as
\begin{equation}
\vec{p}(m) = \begin{pmatrix} p_R(m) \\ p_E(m) \\ p_P(m) \end{pmatrix},\
\mathbf{K} = \begin{pmatrix} -k_R^- & k_R^+ & 0 \\
                        k_R^- & -k_R^+ -k_P^+ & k_P^- \\
                        0 & k_P^+ & -k_P^- 
                \end{pmatrix},
\label{eq:3state_cme_matrices_pt1}
\end{equation}
and the initiation matrices given by
\begin{equation}
\mathbf{R_A} = \begin{pmatrix}
                0 & 0 & 0 \\ 
                0 & 0 & r \\ 
                0 & 0 & 0
                \end{pmatrix},\
\mathbf{R_D} = \begin{pmatrix}
                0 & 0 & 0 \\ 
                0 & 0 & 0 \\ 
                0 & 0 & r
                \end{pmatrix}.
\label{eq:3state_cme_matrices_pt2}
\end{equation}
The elements of $\vec{p}(m)$ encode the probabilities of having $m$ mRNA present
along with the promoter having repressor bound ($R$), being empty ($E$), or
having RNAP bound ($P$), respectively. $\mathbf{R_A}$ describes probability flux
\textit{arriving} at the state $\vec{p}(m)$ from a state with one fewer mRNA,
namely $\vec{p}(m-1)$, and $\mathbf{R_D}$ describes probability flux
\textit{departing} from the state $\vec{p}(m)$ for a state with one more mRNA,
namely $\vec{p}(m+1)$. $\mathbf{K}$ is closely analogous to model 1.

Mean mRNA at steady state is found analogously to model 1, with the result
\begin{equation}
\langle m\rangle = \frac{r}{\gamma}
        \frac{k_R^- k_P^+}
        {k_R^- k_P^+ + k_R^- (k_P^- + r) + k_R^+ (k_P^- + r)},
\label{eq:model2_meanm}
\end{equation}
and with details again deferred to Appendix~\ref{sec:non_bursty}.
Fold-change is again found from the ratio prescribed by Eq.~\ref{eq:fc_def}, from
which we have
\begin{align}
\text{fold-change}
&=      \frac{k_R^- k_P^+}
        {k_R^- k_P^+ + k_R^- (k_P^- + r) + k_R^+ (k_P^- + r)}
        \frac{k_P^+ + k_P^- + r}{k_P^+}
\\
&=      \left(1 + \frac{k_R^+}{k_R^-}
                \frac{k_P^- + r}{k_P^+ + k_P^- + r}
        \right)^{-1}
\\
&=      \left(1 + \frac{k_R^+}{k_R^-}
        \left(1 + \frac{k_P^+}{k_P^- + r}\right)^{-1}
        \right)^{-1},
\end{align}
which follows the master curve of~\fig{fig1:means_cartoons}(D) with $\rho = 1 +
k_P^+/(k_P^- + r)$ as claimed.

\subsubsection{Nonequilibrium Model Three - Multistep Transcription Initiation
and Escape}
One might reasonably complain that the first two ``nonequilibrium'' models we
have considered are straw men. Their steady states necessarily satisfy detailed
balance which is equivalent to thermodynamic equilibrium. Why is this the case?
At steady state there is by definition no net probability flux in or out of each
promoter state, but since the promoter states form a linear chain, there is only
one way in or out of the repressor bound and RNAP bound states, implying each
edge must actually have a net zero probability flux, which is the definition of
detailed balance (usually phrased as equality of forward and reverse transition
fluxes).

Now we consider model 3 in Figure~\ref{fig1:means_cartoons}(C) which allows the
possibility of true nonequilibrium steady-state fluxes through the promoter
states. We point out that this model was considered previously
in~\cite{Mitarai2015} where a comparison was made with model 1 as used
in~\cite{Jones2014}. The authors of~\cite{Mitarai2015} argued that the
additional complexity is essential to properly account for the noise in the mRNA
distribution. We will weigh in on both models later when we consider observables
beyond fold-change.

The master equation governing this model is identical in form to model 2 above,
namely
\begin{equation}
\deriv{t}\vec{p}(m,t) =
\left( \mathbf{K} - \mathbf{R_D} - \gamma m \mathbf{I} \right) \vec{p}(m,t)
                + \mathbf{R_A} \vec{p}(m-1,t) +
                \gamma (m+1) \mathbf{I} \vec{p}(m+1,t),
\end{equation}
but with a higher-dimensional state space and different matrices. The
state vector and promoter transition matrix are now
\begin{equation}
\vec{p}(m) = \begin{pmatrix} p_R(m) \\ p_E(m) \\
                             p_C(m) \\ p_O(m)\end{pmatrix},\
\mathbf{K} = \begin{pmatrix} -k_R^- & k_R^+ & 0 & 0\\
                        k_R^- & -k_R^+ -k_P^+ & k_P^- & 0 \\
                        0 & k_P^+ & -k_P^- - k_O & 0 \\
                        0 & 0 & k_O & 0
                \end{pmatrix},
\end{equation}
with the four promoter states, in order, being repressor bound ($R$), empty
($E$), RNAP closed complex ($C$), and RNAP open complex ($O$). Besides
increasing dimension by one, the only new feature in $\mathbf{K}$ is the
rate $k_O$, representing the rate of open complex formation from
the closed complex, which we assume for simplicity to be irreversible
in keeping with some~\cite{Mitarai2015} but not all~\cite{DeHaseth1998}
past literature. The initiation matrices are given by
\begin{equation}
\mathbf{R_A} = \begin{pmatrix}
        0 & 0 & 0 & 0 \\ 
        0 & 0 & 0 & r \\ 
        0 & 0 & 0 & 0 \\ 
        0 & 0 & 0 & 0
                \end{pmatrix},\
\mathbf{R_D} = \begin{pmatrix}
        0 & 0 & 0 & 0 \\ 
        0 & 0 & 0 & 0 \\ 
        0 & 0 & 0 & 0 \\ 
        0 & 0 & 0 & r
                \end{pmatrix},
\end{equation}
again closely analogous to nonequilibrium model 2.

The expression for mean mRNA is substantially more complicated now, as worked
out in Appendix~\ref{sec:non_bursty} where we find
\begin{equation}
\langle m\rangle = \frac{r}{\gamma}
        \frac{k_R^- k_P^+ k_O}
        {k_R^- [(k_P^+ (k_O + r) + r(k_P^- + k_O)] + k_R^+ r(k_P^- + k_O)},
\label{eq:model3_mean_m}
\end{equation}
which can be simplified to
\begin{equation}
\langle m\rangle
= \frac{r}{\gamma}
\frac{\frac{k_P^+ k_O}{r(k_O + k_P^-)}}
        {1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)} + \frac{k_R^+}{k_R^-}}.
\end{equation}
The strategy is to isolate the terms involving the repressor, so that now the
fold-change is seen to be simply
\begin{align}
\text{fold-change}
&= \frac{\frac{k_P^+ k_O}{r(k_O + k_P^-)}}
        {1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)} + \frac{k_R^+}{k_R^-}}
        \frac{1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)}}
                {\frac{k_P^+ k_O}{r(k_O + k_P^-)}}
\\
&= \left(
        1 + \frac{k_R^+}{k_R^-}
        \left(1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)}\right)^{-1}
\right)^{-1},
\end{align}
surprisingly reducing to the master curve of~\fig{fig1:means_cartoons}(D) once
again, with $\rho = 1 + \frac{k_P^+ (k_O + r)}{r(k_O + k_P^-)}$.

This example hints that an arbitrarily fine-grained model of downstream
transcription steps may still be collapsed to the form of the master curve for
the means given in Figure~\ref{fig1:means_cartoons}(D), so long as the repressor
binding is exclusive with transcriptionally active states. We offer this as a
conjecture, and we suspect that a careful argument using the King-Altman diagram
method~\cite{King1956, Hill1966} might furnish a ``proof.'' Our focus here is
not on full generality but rather to survey an assortment of plausible models
for simple repression  that have been proposed in the literature.

\subsubsection{Nonequilibrium Model Four - ``Active'' and ``Inactive'' States}
Model 4 in Figure~\ref{fig1:means_cartoons}(C) is at the core of the theory
in~\cite{Razo-Mejia2020}. At a glance the cartoon for this model may appear very
similar to model 2, and mathematically it is, but the interpretation is rather
different. In model 2, we interpreted the third state literally as an RNAP-bound
promoter and modeled initiation of a transcript as triggering a promoter state
change, making the assumption that an RNAP can only
make one transcript at a time. In contrast, in the present model the promoter
state does \textit{not} change when a transcript is initiated. So we no longer
interpret these states as literally RNAP bound and unbound but instead as
coarse-grained ``active'' and ``inactive'' states, the details of which we leave
unspecified for now. We will comment more on this model below when we discuss
Fano factors of models.

Mathematically this model is very similar to models 1 and 2. Like model 1, the
matrix $R$ describing transcription initiation is diagonal, namely
\begin{equation}
\mathbf{R} = \begin{pmatrix}
                0 & 0 & 0 \\ 
                0 & 0 & 0 \\ 
                0 & 0 & r
        \end{pmatrix}.
\end{equation}
The master equation takes verbatim the same form as it did for model 1,
Eq.~\ref{eq:2state_rep_cme}. Meanwhile the promoter transition
matrix $\mathbf{K}$ is the same as Eq.~\ref{eq:3state_cme_matrices_pt1}
from model 2, although we relabel the
rate constants from $k_P^\pm$ to $k^\pm$ to reiterate that these are not simply
RNAP binding and unbinding rates.

Carrying out the algebra, the mean mRNA can be found to be
\begin{equation}
\langle m\rangle = \frac{r}{\gamma}
\frac{k_R^- k^+}
{k_R^- k^+ + k_R^- k^- + k_R^+ k^-},
\end{equation}
and the fold-change readily follows,
\begin{align}
\text{fold-change}
&=      \frac{k_R^- k^+}{k_R^- k^+ + k_R^- k^- + k_R^+ k^-}
        \frac{k_R^- k^+ + k_R^- k^-}{k_R^- k^+}
\\
&=      \left(1 + \frac{k_R^+}{k_R^-}
                \left(1 + \frac{k^+}{k^-}\right)^{-1}
        \right)^{-1},
\end{align}
from which we see $\rho = 1 + k^+/k^-$ as shown in~\fig{fig1:means_cartoons}(C).

\subsubsection{Nonequilibrium Model Five - Bursty Promoter}
The final model we consider shown in Figure~\ref{fig1:means_cartoons}(C) is an
intuitive analog to model 1, with just two states, repressor bound or unbound,
and transition rates between them of $k_R^+$ and $k_R^-$. In model 1, when in
the unbound state, single mRNA transcripts are produced as a Poisson process
with some characteristic rate $r$. The current model by contrast produces, at
some Poisson rate $k_i$, \textit{bursts} of mRNA transcripts.
The burst sizes are assumed to be geometrically distributed with
a mean burst size $b$, which we will motivate in Section~\ref{sec:beyond_means}
when we derive this model as a certain limiting case of model 4.

From this intuitive picture and by analogy to model 1, then, it should be
plausible that the mean mRNA level is
\begin{equation}
\langle m\rangle = \frac{k_i b}{\gamma}
        \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1},
\end{equation}
which will turn out to be correct from a careful calculation. For now, we simply
note that just like model 1, the fold-change becomes
\begin{equation}
\text{fold-change} = \left(1 + \frac{k_R^+}{k_R^-}\right)^{-1}
\end{equation}
with $\rho=1$ also like model 1.
We will also see later how this model emerges as a natural limit of model 4.
\mrm{ENDS HERE}

@@

\begin{figure}[h!]
\centering
\includegraphics{../../figures/si/figS0X_two_state.pdf}
\caption{
\textbf{Two-state promoter chemical master equation.}
(A) Schematic of the two state promoter simple repression model. Rates $k_R^+$
and $k_R^-$ are the association and dissociation rates of the transcriptional
repressor, respectively, $r$ is the transcription initiation rate, and $\gamma$
is the mRNA degradation rate. (B) Schematic depiction of the mRNA count state
transitions. The model in (A) only allows for jumps in mRNA of size 1. The
production of mRNA can only occur when the promoter is in the transcriptionally
active state.}
\label{fig:two_state}
\end{figure}
